Efficient Computation of Extended Sources

### 2021-05-31 repro/papers/sseprecond/paper.tex
interim

The analysis also reveals that the solution components not continuous
at $z=z_s$ are odd there:
\begin{equation}
  \label{eqn:odd1}
  \lim_{z\rightarrow z_s^+} v^{\pm}_{z} \approx - \lim_{z\rightarrow z_s^-}
  v^{\pm}_{z}
\end{equation}
for the solution of \ref{eqn:awepm} with $f_s=0$.
Similarly, 
\begin{equation}
  \label{eqn:odd2}
  \lim_{z\rightarrow z_s^+} p^{\pm}\approx - \lim_{z\rightarrow z_s^-}
  p^{\pm}
\end{equation}
for the solution of \ref{eqn:awepm} with $h_s=0$. Here ``$\approx$''
means in the sense of high frequency asymptotics, that is, that the
difference between the two sides is relatively smooth, hence small if
the data is highly oscillatory. Therefore if $f_s=0$ in system \ref{eqn:awepm},
\begin{equation}
  h_s = \Lambda^{\pm}_{z_s}P_sp^{\pm} = -[v^{\pm}_{z}]|_{z=z_s} \approx -2
  P_sv^{\pm}_{z}
  \label{eqn:tracejump10}
\end{equation}
Similarly, if $h_s=0$ in system \ref{eqn:awepm}, then
\begin{equation}
  \label{eqn:tracejump20}
  f_s = -[p^{\pm}]|_{z=z_s} \approx -2 P_s p^{\pm}.
\end{equation}
Thus $f_s$ determines approximately the boundary value of $p^{\pm}$,
as a solution of the acoustic wave system in the half-space
$z>z_s$. However, as repeated in equation \ref{eqn:tracejump10}, a
solution with this boundary value is also the restriction to $z>z_s$
of a solution to \ref{eqn:awepm} with $f_s=0$ and $h_s=
\Lambda^{\pm}_{z_s}P_sp^{\pm}$. Therefore if
\begin{equation}
  \label{eqn:hfcondn}
  h_s =-\frac{1}{2}\Lambda^{\pm}_{z_s}f_s,
\end{equation}
then the pressure boundary value $P_sp^{\pm}$ is the
same for the solutions of \ref{eqn:awepm} for source vectors $(h_s,0)$
and $(0,f_s)$. Since the pressure boundary values are the same, the solutions
in $z>z_s$ are the same. In particular, since $z_r>z_s$ and ${\cal
  S}^{\pm}_{z_s,z_r}(h_s,f_s)^T = (P_rp^{\pm},P_rv^{\pm}_z)^T$, it follows
that
\begin{equation}
  \label{eqn:snull}
  {\cal S}^{\pm}_{z_s,z_r}\left(-\frac{1}{2}\Lambda^{\pm}_{z_s}f_s,f_s\right)^T \approx 0.
\end{equation}

@@

These source functions satisfy
the relation \ref{eqn:hfcondn}, therefore source vectors $(h_s,0)$ and
$(0, f_s)$ generate {\em the same acoustic field} $(p^+,\bv^+)$ in $z>z_s$. Figures
\ref{fig:drecphh0}, \ref{fig:dfwdphh0}, \ref{fig:daltphh0} (homogeneous model) and
\ref{fig:drecplh0}, \ref{fig:dfwdplh0}, \ref{fig:daltplh0} (lens model) show the pressure
gathers extracted at $z_r=1000$ m for the point source at $z=z_d$ and
for the two choices of extended source at $z=z_s$, on the same color
scale. The obvious similarity between the fields generated by the two
extended sources,
predicted by equation \ref{eqn:snull}, is confirmed by trace
comparisons in figures \ref{fig:drecphh0tr21}-\ref{fig:daltplh0tr81}.

\section{Approximate Inversion}

Computation of $\Lambda^{\pm}_{z_s}$ will prove a critical step in an
efficient iterative solution of the linear subproblem
\ref{eqn:esis}. Direct computation, for instance by solving
\ref{eqn:sawe2} and reading off $P_sv^{\pm}_{2,z}$, turns out to be
numerically ill-behaved. The relation \ref{eqn:snull} provides and
alternative approach, {\em provided} that an accurate approximate
inverse to $S^+_{z_s,z_r}$ is available. For instance, the first row
of the vector equation \ref{eqn:snull} reads (slightly rearranged):
\[
\Pi_0{\cal S}^+_{z_s,z_r}\Pi_1^T f_s \approx \frac{1}{2}\Pi_0{\cal
  S}^+_{z_s,z_r}\Pi_0\Lambda^+_{z_s}f_s
\]
\begin{equation}
  \label{eqn:lamidea}
  = \frac{1}{2} S^+_{z_s,z_r}\Lambda^+_{z_s}f_s.
\end{equation}
An approximate inverse for $S^++_{z_s,z_r}$ would permit (approximate)
solution of this equation for $\Lambda^+_{z_s}f_s$.

@@

I have applied this procedure to the pressure gathers shown in Figures
\ref{fig:recphh0} and \ref{fig:recplh0}, generated by a point source
at $z_s=3000$ m, $x_s=3500$ m, propagating in the homogeneous and lens
models respectively. The source time dependence is again a
[1, 2.5, 10, 12.5] Hz trapezoidal bandpass filter. As noted earlier,
a point source at depth $z=z_s$ generates a non-downgoing acoustic
field in the slab $z_s < z < z_r$. However the pressure gather at
$z=z_r$ contains only downgoing energy after a suitable mute: this is
evident in the figures. Therefore one would expect inversion of this
data to produce an aperture-limited (in effect, dip-filtered) version
of the point source.

Figure \ref{fig:srcvzghh0} shows the approximate inversion suggested
by the first equation in \ref{eqn:approxinv}. The result indeed looks
like a dip-filtered point source. To see that this is indeed an
inversion, I apply $S^{+}_{z_s,z_r}$ to the source shown in Figure
\ref{fig:srcvzghh0}, to produce the re-simulation shown in Figure
\ref{fig:fwdphh0}. Comparison with the gather in Figure
\ref{fig:recphh0} via a difference plot \ref{fig:diffrecphh0} verifies
that a very close match has been obtained, so the source in Figure
\ref{fig:srcvzghh0} is indeed a satisfactory inversion, even though
the original point source has not been recovered: the difference is in
the null space of the modeling operator, arising from the limited
aperture of the acquisition geometry.

As mentioned at the beginning of this section, the availability of approximate two-sided inverses for $S^{+},
V^{+}$ leads to a practical computation for $\Lambda^{+}$: apply
$4(V^{+}_{z_s,z_r})^T$ to both sides of equation \ref{eqn:lamidea} and
use the first equation in the list \ref{eqn:approxinv} to get
\begin{equation}
  \label{eqn:lamident}
  \Lambda^+_{z_s} \approx 8(V^{+}_{z_s,z_r})^T\Pi_0 {\cal S}^+_{z_s,z_r}\Pi_1^T.
\end{equation}
This identity is the second major result of this section: it shows how
to compute that action of $\Lambda^+_{z_s}$ by propagating the input
pressure trace, identified as a source for the velocity evolution,
forward in time from $z_s$ to $z_r$
reading off the pressure trace on $z=z_r$, identifying it once more as
a point load (source for velocity), propagating it backwards in time from $z_r$ to
$z_s$, and finally reading off the velocity trace, interpreted as a
pressure evolution source on $z_s$.

@@

Computation of the transpose of $\Lambda^+$ (exact, not approximate in
the high frequency sense) is critical to the successful
construction of the preconditioner. The relation \ref{eqn:lamident}
does not provide a computation for this operator. However set
\begin{equation}
  \label{eqn:lamtilde}
  \tilde{\Lambda}^+_{z_s} = 8(V^{+}_{z_s,z_r})^T\Pi_0 {\cal S}^+_{z_s,z_r}\Pi_1^T.
\end{equation}
Then \ref{eqn:lamident} can be rewritten
\[
  \Lambda^+_{z_s} \approx \tilde{\Lambda}^+_{z_s}.
\]
Since successful preconditioning requires only approximate inversion,
use of $\tilde{\Lambda}^+_{z_s} $ in place of $\Lambda^+_{z_s}$ will
still yield a working preconditioner, and the former can be transposed
to machine precision via the definition \ref{eqn:lamtilde} and the adjoint state
method (equations \ref{eqn:trtr} \ref{eqn:trtrcomp}):
\begin{equation}
  \label{eqn:lamtransp}
   (\tilde{\Lambda}^+_{z_s})^T = 8 \Pi_1 {\cal S}^-_{z_r,z_s}\Pi_0^T V^{+}_{z_s,z_r}
\end{equation}

@@

\plot{dsrcpghh0}{width=\textwidth}{Trace $P_sp^+$ on $z=z_s=3000$ m of
  pressure field from point source at $z_d=3500$ m, $x_d=3500$ m,
  bandpass filter source. Scaled to represent point load (velocity
  source). }

\plot{dsrcvzghh0}{width=\textwidth}{$\Lambda^+_{z_s}P_sp^+$ = scaled
  version of trace $P_sv_z^+$ on $z=z_s=3000$ m of
  vertical velocity field from point source at $z_d=3500$ m, $x_d=3500$ m,
  bandpass filter source.}

@@

\plot{dfwdphh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the homogeneous model described in the 
  text, with extended pressure source (constitutive defect) $h_s$ on $z=z_s=3000$ m given by  field depicted in Figure \ref{fig:dsrcvzghh0}
  ($h_s=-2P_sv_z^+=\Lambda^+_{z_s}P_sp^+$) and zero velocity source (vertical load) 
  ($f_s=0$).}

\plot{daltphh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the homogeneous model described in the 
  text, with extended velocity source (vertical load) $f_s$ on $z=z_s=3000$
  m given by the field depicted in Figure \ref{fig:dsrcpghh0}
  ($f_s=-2P_sp^+$) and zero pressure source (constitutive defect)
  ($h_s=0$).}

@@

\plot{dfwdplh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the lens model described in the 
  text, with extended pressure source (constitutive defect) on $z=z_s=3000$ m given bythe field depicted in Figure \ref{fig:dsrcvzghh0}
   ($h_s=-2P_sv_z^+=\Lambda^+_{z_s}P_sp^+$) and zero velocity source (vertical load) 
  ($f_s=0$).}

\plot{daltplh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the lens model described in the 
  text, with extended velocity source (vertical load) on $z=z_s=3000$
  m given by the field depicted in Figure \ref{fig:dsrcpghh0}
  ($f_s=-2P_sp^+$) and zero pressure source (constitutive defect)
  ($h_s=0$).}

@@

@@

\plot{recphh0}{width=\textwidth}{Pressure gather from point source at
  $z_s=3000$ m, $x_s=3500$ m, homogeneous model. Mute applied near cable ends.}

\plot{srcvzghh0}{width=\textwidth}{Approximate inversion of the gather
  in Figure \ref{fig:recphh0} suggested by the first equation in
  display \ref{eqn:approxinv}. Scaled version of output $v_z$ field
  obtained by applying transpose of $V^+_{z_s,z_r}$.}

\plot{fwdphh0}{width=\textwidth}{Pressure gather produced from source
  shown in Figure \ref{fig:srcvzghh0}.}

\plot{diffrecphh0}{width=\textwidth}{Difference between gathers
  displayed in Figures \ref{fig:recphh0} and \ref{fig:fwdphh0},
  plotted on same color scale.}

\plot{invvzghh0}{width=\textwidth}{Inverted pressure source (= scaled
  $v_z$) gather obtained by another application of $(V^+_{z_s,z_r})^T$
  to data shown in Figure \ref{fig:fwdphh0}. Compare Figure
  \ref{fig:srcvzghh0}.}

>>>>

The analysis also reveals that the solution components not continuous
at $z=z_s$ are odd there:
\begin{equation}
  \label{eqn:odd1}
  \lim_{z\rightarrow z_s^+} v^{\pm}_{z} \approx - \lim_{z\rightarrow z_s^-}
  v^{\pm}_{z}
\end{equation}
for the solution of \ref{eqn:awepm} with $f_s=0$.
Similarly, 
\begin{equation}
  \label{eqn:odd2}
  \lim_{z\rightarrow z_s^+} p^{\pm}\approx - \lim_{z\rightarrow z_s^-}
  p^{\pm}
\end{equation}
for the solution of \ref{eqn:awepm} with $h_s=0$. Here ``$\approx$''
means in the sense of high frequency asymptotics, that is, that the
difference between the two sides is relatively smooth, hence small if
the data is highly oscillatory. Therefore if $f_s=0$ in system \ref{eqn:awepm},
\begin{equation}
  h_s = \Lambda^{\pm}_{z_s}P_sp^{\pm} = -[v^{\pm}_{z}]|_{z=z_s} \approx -2
  P_sv^{\pm}_{z}
  \label{eqn:tracejump10}
\end{equation}
Similarly, if $h_s=0$ in system \ref{eqn:awepm}, then
\begin{equation}
  \label{eqn:tracejump20}
  f_s = -[p^{\pm}]|_{z=z_s} \approx -2 P_s p^{\pm}.
\end{equation}
Thus $f_s$ determines approximately the boundary value of $p^{\pm}$,
as a solution of the acoustic wave system in the half-space
$z>z_s$. However, as repeated in equation \ref{eqn:tracejump10}, a
solution with this boundary value is also the restriction to $z>z_s$
of a solution to \ref{eqn:awepm} with $f_s=0$ and $h_s=
\Lambda^{\pm}_{z_s}P_sp^{\pm}$. Therefore if
\begin{equation}
  \label{eqn:hfcondn}
  h_s =-\frac{1}{2}\Lambda^{\pm}_{z_s}f_s,
\end{equation}
then the pressure boundary value $P_sp^{\pm}$ is the
same for the solutions of \ref{eqn:awepm} for source vectors $(h_s,0)$
and $(0,f_s)$. Since the pressure boundary values are the same, the solutions
in $z>z_s$ are the same. In particular, since $z_r>z_s$ and ${\cal
  S}^{\pm}_{z_s,z_r}(h_s,f_s)^T = (P_rp^{\pm},P_rv^{\pm}_z)^T$, it follows
that
\begin{equation}
  \label{eqn:snull}
  {\cal S}^{\pm}_{z_s,z_r}\left(\frac{1}{2}\Lambda^{\pm}_{z_s}f_s,f_s\right)^T \approx 0.
\end{equation}

@@

These source functions satisfy
the relation \ref{eqn:hfcondn}, therefore source vectors $(h_s,0)$ and
$(0, f_s)$ generate {\em the same acoustic field} $(p^+,\bv^+)$ in $z>z_s$. Figures
\ref{fig:drecphh0}, \ref{fig:dfwdphh0}, \ref{fig:daltphh0} (homogeneous model) and
\ref{fig:drecplh0}, \ref{fig:dfwdplh0}, \ref{fig:daltplh0} (lens model) show the pressure
gathers extracted at $z_r=1000$ m for the point source at $z=z_d$ and
for the two choices of extended source at $z=z_s$, on the same color
scale. The obvious similarity between the fields generated by the two
extended sources,
predicted by equation \ref{eqn:snull}, is confirmed by trace
comparisons in figures \ref{fig:drecphh0tr21}-\ref{fig:daltphh0tr81}
and
\ref{fig:drecplh0tr21}-\ref{fig:daltplh0tr81}.

\section{Time Reversal}

@@

I have applied this procedure to the pressure gathers shown in Figures
\ref{fig:drecphh0} and \ref{fig:drecplh0}, generated by a point source
at $z_d=3500$ m, $x_s=3500$ m, propagating in the homogeneous and lens
models respectively. The source time dependence is again a [1, 2.5,
10, 12.5] Hz trapezoidal bandpass filter. As noted earlier, in the
region $z < z_s=3000 m$, these fields are downgoing and can be
regarded as the result of either pressure or velocity source at $z=z_s$: the
pressure source $h_s$ (Figure \ref{fig:dhshh0}) is -2 times the
vertical velocity gather depicted in Figure \ref{fig:dsrcvzhh0}, the
velocity source $f_s$ (Figure \ref{fig:dfshh0}) is -2 times the
pressure gather depicted in Figure \ref{fig:dsrcphh0}.

Figure \ref{fig:dinvhshh0} shows the approximate inversion (via the
first equation in display \ref{eqn:approxinv}) of the
pressure gather shown in Figure \ref{fig:drecphh0}, inverted in the homogeneous model
used to generate the data. The result is a scaled, dip-filtered
version of the pressure source shown in Figure \ref{fig:dhshh0}.
The dip filter effect is unavoidable and is caused by the aperture
limitation of the acquisition geometry: the steeper dips in the source
gather \ref{fig:dhshh0} do not contribute to the data, and are not
recovered in the inversion.

To see that this is indeed an inversion, I apply $S^{+}_{z_s,z_r}$ to
the source shown in Figure \ref{fig:dinvhshh0}, to produce the
re-simulation shown in Figure \ref{fig:drerecphh0}. Simulation is carried
out in the homogeneous model used in the inversion. Comparison with the
gather in Figure \ref{fig:drecphh0} via a difference plot
\ref{fig:ddiffrecphh0} verifies that a very close match has been
obtained, so the source in Figure \ref{fig:dinvhshh0} is indeed a
satisfactory inversion, even it differs considerably from the original
point source: the difference lies in an approximate null space of the modeling
operator, arising from the limited aperture of the acquisition
geometry.

The remainder of this section is devoted to illustrating two important
features of the approximate inversion developed above. The
first is the validity of the inversion approach given in display
\ref{eqn:approxinv} so long as the fields generated by the
model used in simulation and inversion are downgoing. The models
may be quite far from homogeneous, and even generate triplications,
whereas the inversions remain accurate modulo the dip filtering effect
of aperture. Figure \ref{fig:dinvhsll0} shows the pressure source at $z=z_s$
inverted from the data Figure
\ref{fig:drecpll0}, in which a triplication is obvious. Recall that
this data is generated by propagating a point source at $z_d=3500$ m
in the lens model. The inversion uses the same model. Figure
\ref{fig:dinvhsll0} is an aperture-limited version of the pressure
source at $z=z_s=3000$ m
(Figure \ref{fig:dhsll0}) that generates this data by a dip filter, in
the same way that the sources in Figures \ref{fig:dhsll0} and
\ref{fig:dinvhsll0} are related. The inverted source is an accurate
inversion: it generates the data depicted in Figure
\ref{fig:drerecpll0}, which is very close to the input data (Figure
\ref{fig:drecpll0}). The difference is displayed on the same color
scale in Figure \ref{ddiffrecpll0}.

The second feature is a decoupling of the models used (implicitly or
otherwise) in simulation and those used in inversion: it is possible
to {\em accurately invert pressure data for pressure source using the
  wrong model}, in the sense that the inverted source will
generate an accurate recovery of the input data re-simulated in the
same (wrong) model. For example, inversion of the data of the previous
example (Figure \ref{fig:drecpll0}) {\em in the homogeneous model}
results in an inverted source gather at $z=z_s$ (Figure \ref{fig:dinvhslh0}
that is very far from a dip-filtered version of the source used
generate the data \ref{eqn:dhshh0}. However, this result is an
accurate inversion: re-simulation {\em using the same (homogeneous)
  model as used in the inversion} results in accurate recovery of the
input pressure gather (Figure \ref{fig:drecpll0}). The difference is
shown on the same color scale in Figure \ref{fig:ddiffrecplh0}.

\section{Computing $\Lambda$}

Computation of $\Lambda^{\pm}_{z_s}$ will prove a critical step in an
efficient iterative solution of the linear subproblem
\ref{eqn:esis}. Direct computation of the pressure-to-source operator $\Lambda^{\pm}_{z_s}$, for instance by solving
\ref{eqn:awepm} and reading off $P_sv^{\pm}_{z}$, turns out to be
numerically ill-behaved. The relation \ref{eqn:snull} provides and
alternative approach, taking advantage of the accurate approximate
inverse to $S^+_{z_s,z_r}$ is available. For instance, the first row
of the vector equation \ref{eqn:snull} reads (slightly rearranged):
\[
\Pi_0{\cal S}^+_{z_s,z_r}\Pi_1^T f_s \approx -\frac{1}{2}\Pi_0{\cal
  S}^+_{z_s,z_r}\Pi_0\Lambda^+_{z_s}f_s
\]
\begin{equation}
  \label{eqn:lamidea}
  = -\frac{1}{2} S^+_{z_s,z_r}\Lambda^+_{z_s}f_s.
\end{equation}

The approximate inverse construction for $S^++_{z_s,z_r}$ would permits (approximate)
solution of this equation for $\Lambda^+_{z_s}f_s$: apply
$4(V^{+}_{z_s,z_r})^T$ to both sides of equation \ref{eqn:lamidea} and
use the first equation in the list \ref{eqn:approxinv} to get
\begin{equation}
  \label{eqn:lamident}
  \Lambda^+_{z_s} \approx -8(V^{+}_{z_s,z_r})^T\Pi_0 {\cal S}^+_{z_s,z_r}\Pi_1^T.
\end{equation}
This identity is the second major result of this section: it shows how
to compute that action of $\Lambda^+_{z_s}$ by propagating the input
pressure trace, identified as a source for the velocity evolution,
forward in time from $z_s$ to $z_r$
reading off the pressure trace on $z=z_r$, identifying it once more as
a point load (source for velocity), propagating it backwards in time from $z_r$ to
$z_s$, and finally reading off the velocity trace, interpreted as a
pressure evolution source on $z_s$.

@@

Figure \ref{fig:preddinvhshh0} shows the pressure source gather produced
from the pressure data gather in Figure \ref{drecphh0} by application
of the operator on the right-hand side of
formula \ref{eqn:lamident}. The scale is the same as for Figures
\ref{fig:dinvhshh0} and \ref{fig:dhshh0}. In fact the former
appears nearly identical, and both appear to be dip-filtered versions
of the latter. The difference plot (Figure \ref{fig:ddiffinvhshh0},
using the same color scale, shows that the similarity is quantitative.

Figures \ref{fig:preddinvhsll0} and \ref{fig:ddiffinvhsll0} accomplish
the same comparison for the $\Lambda^+_{z_s}$ approximation in
equation \ref{eqn:lamident}. A deeper study of the pressure-to-source
operator shows that it is approximately dependent only on the model
coefficients near the source surface ($z=z_s$ in this case). Since the
homogeneous and lens models are identical near this surface, it is
unsurprising that these figures are very close to the previous two.

Computation of the transpose of $\Lambda^+$ (exact, not approximate in
the high frequency sense) is critical to the successful
construction of the preconditioner. The relation \ref{eqn:lamident}
does not provide a computation for this operator. However set
\begin{equation}
  \label{eqn:lamtilde}
  \tilde{\Lambda}^+_{z_s} = -8(V^{+}_{z_s,z_r})^T\Pi_0 {\cal S}^+_{z_s,z_r}\Pi_1^T.
\end{equation}
Then \ref{eqn:lamident} can be rewritten
\[
  \Lambda^+_{z_s} \approx \tilde{\Lambda}^+_{z_s}.
\]
Since successful preconditioning requires only approximate inversion,
use of $\tilde{\Lambda}^+_{z_s} $ in place of $\Lambda^+_{z_s}$ will
still yield a working preconditioner, and the former can be transposed
to machine precision via the definition \ref{eqn:lamtilde} and the adjoint state
method (equations \ref{eqn:trtr} \ref{eqn:trtrcomp}):
\begin{equation}
  \label{eqn:lamtransp}
   (\tilde{\Lambda}^+_{z_s})^T = -8 \Pi_1 ({\cal S}^+_{z_r,z_s})^T\Pi_0^T V^{+}_{z_s,z_r}
 \end{equation}

\section{Unitarity}

@@

\plot{dsrcphh0}{width=\textwidth}{Trace $P_sp^+$ on $z=z_s=3000$ m of
  pressure field from point source at $z_d=3500$ m, $x_d=3500$ m,
  bandpass filter source. }

\plot{dsrcvzhh0}{width=\textwidth}{Trace $P_sv_z^+$ on $z=z_s=3000$ m of
  vertical velocity field from point source at $z_d=3500$ m, $x_d=3500$ m,
  bandpass filter source.}

@@

\plot{dfwdphh0}{width=\textwidth}{Pressure gather at receiver depth
  $z_r=1000$ m from field generated by causal solution of acoustic
  system \ref{eqn:awepm} in the homogeneous model described in the
  text, with extended pressure source (constitutive defect)
  on $z=z_s=3000$ m given by $h_s=-2P_sv_z^+=\Lambda^+_{z_s}P_sp^+$,
  where $P_sp^+$ is shown in Figure \ref{fig:dsrcphh0} and $P_sv^+_z$
  in Figure \ref{fig:dsrcvzhh0}. In this simulation, the velocity
  velocity source (vertical load) vanishes: $f_s=0$.}

\plot{daltphh0}{width=\textwidth}{Pressure gather at receiver depth
  $z_r=1000$ m from field generated by causal solution of acoustic
  system \ref{eqn:awepm} in the homogeneous model described in the
  text, with extended velocity source (vertical load) $f_s=-2P_sp^+$
  on $z=z_s=3000$ m, where $P_sp^+$ is shown in Figure
  \ref{fig:dsrcphh0}.  In this simulation, the zero pressure source
  (constitutive defect vanishes: $h_s=0$.}

@@

@@

\plot{daltphh0tr21}{width=\textwidth}{Overplot of traces 21 ($x=2400$) 
  from gathers shown in \ref{fig:drecphh0} (blue), \ref{fig:daltphh0}
  (red).}

\plot{drecphh0tr51}{width=\textwidth}{Overplot of traces 51 ($x=3000$) 
  from gathers shown in \ref{fig:drecphh0} (blue), \ref{fig:dfwdphh0}
  (red).}

\plot{daltphh0tr51}{width=\textwidth}{Overplot of traces 51 ($x=3000$) 
  from gathers shown in \ref{fig:drecphh0} (blue), \ref{fig:daltphh0}
  (red).}

\plot{drecphh0tr81}{width=\textwidth}{Overplot of traces 81 ($x=3600$) 
  from gathers shown in \ref{fig:drecphh0} (blue), \ref{fig:dfwdphh0}
  (red).}

\plot{daltphh0tr81}{width=\textwidth}{Overplot of traces 81 ($x=3600$) 
  from gathers shown in \ref{fig:drecphh0} (blue), \ref{fig:daltphh0}
  (red).}

\plot{dfwdplh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the lens model described in the 
  text, with extended pressure source (constitutive defect) on
  $z=z_s=3000$ m given bythe field depicted in Figure
  \ref{fig:dsrcvzhh0} scaled by -2 
   ($h_s=-2P_sv_z^+=\Lambda^+_{z_s}P_sp^+$) and zero velocity source (vertical load) 
  ($f_s=0$).}

\plot{daltplh0}{width=\textwidth}{Pressure gather at receiver depth 
  $z_r=1000$ m from field generated by causal solution of acoustic 
  system \ref{eqn:awepm} in the lens model described in the 
  text, with extended velocity source (vertical load) on $z=z_s=3000$
  m given by the field depicted in Figure \ref{fig:dsrcphh0} scaled by -2
  ($f_s=-2P_sp^+$) and zero pressure source (constitutive defect)
  ($h_s=0$).}

\plot{drecplh0tr21}{width=\textwidth}{Overplot of traces 21 ($x=2400$) 
  from gathers shown in \ref{fig:drecplh0} (blue), \ref{fig:dfwdplh0}
  (red).}

\plot{daltplh0tr21}{width=\textwidth}{Overplot of traces 21 ($x=2400$) 
  from gathers shown in \ref{fig:drecplh0} (blue), \ref{fig:daltplh0}
  (red).}

\plot{drecplh0tr51}{width=\textwidth}{Overplot of traces 51 ($x=3000$) 
  from gathers shown in \ref{fig:drecplh0} (blue), \ref{fig:dfwdplh0}
  (red).}

\plot{daltplh0tr51}{width=\textwidth}{Overplot of traces 51 ($x=3000$) 
  from gathers shown in \ref{fig:drecplh0} (blue), \ref{fig:daltplh0}
  (red).}

\plot{drecplh0tr81}{width=\textwidth}{Overplot of traces 81 ($x=3600$) 
  from gathers shown in \ref{fig:drecplh0} (blue), \ref{fig:dfwdplh0}
  (red).}

@@

\plot{dinvhshh0}{width=\textwidth}{Approximate inversion via first 
  equation in display \ref{eqn:approxinv}. Inversion in homogenous model of 
  the pressure gather in Figure \ref{fig:drecphh0}, simulated with 
  homogeneous model. Scaled version of output $v_z$ field obtained by 
  applying transpose of $4V^+_{z_s,z_r}$.}

\plot{drerecphh0}{width=\textwidth}{Re-simulated pressure gather
  produced from inverted source
  shown in Figure \ref{fig:dinvhshh0}. Simulation in homogeneous model.}

\plot{ddiffrecphh0}{width=\textwidth}{Difference between gathers
  displayed in Figures \ref{fig:drecphh0} and \ref{fig:drerecphh0},
  plotted on same color scale.}

\plot{dinvhsll0}{width=\textwidth}{Approximate inversion via first 
  equation in display \ref{eqn:approxinv}. Inversion in lens model of 
  the pressure gather in Figure \ref{fig:drecpll0}, simulated with 
  lens model. Scaled version of output $v_z$ field obtained by 
  applying transpose of $4V^+_{z_s,z_r}$.}

\plot{drerecphh0}{width=\textwidth}{Re-simulated pressure gather
  produced from inverted source
  shown in Figure \ref{fig:dinvhsll0}. Simulation in lens model.}

\plot{ddiffrecphh0}{width=\textwidth}{Difference between gathers
  displayed in Figures \ref{fig:drecpll0} and \ref{fig:drerecpll0},
  plotted on same color scale.}

\plot{dinvhslh0}{width=\textwidth}{Approximate inversion via first 
  equation in display \ref{eqn:approxinv}. Inversion in homogeneous model of 
  the pressure gather in Figure \ref{fig:drecplh0}, simulated with 
  lens model. Scaled version of output $v_z$ field obtained by 
  applying transpose of $4V^+_{z_s,z_r}$.}

\plot{drerecplh0}{width=\textwidth}{Re-simulated pressure gather produced from
  inverted source
  shown in Figure \ref{fig:dinvhslh0}. Simulation in homogeneous model.}

\plot{ddiffrecplh0}{width=\textwidth}{Difference between gathers
  displayed in Figures \ref{fig:drecplh0} and \ref{fig:drerecplh0},
  plotted on same color scale.}

\plot{preddinvhshh0}{width=\textwidth}{Pressure source gather = image
  of pressure-to-source operator $\Lambda^+_{z_s}$ applied to pressure gather
  shown in Figure \ref{fig:dsrcphh0}, homogeneous model. Compare Figure
  \ref{fig:dinvhshh0}.}

\plot{ddiffinvhshh0}{width=\textwidth}{Difference between image
  (Figure \ref{fig:predinvhshh0}) of $\Lambda^+_{z_s}$ applied to
  pressure gather (Figure \ref{fig:dsrcphh0}) and source gather
  (Figure \ref{fig:dinvhshh0}) inverted from receiver pressure gather
  (Figure \ref{fig:drecphh0}), homogeneous model used in all
  propagations. Same color scale as in Figure \ref{fig:predinvhshh0}. }

\plot{preddinvhsll0}{width=\textwidth}{Pressure source gather = image
  of pressure-to-source operator $\Lambda^+_{z_s}$ applied to pressure gather
  shown in Figure \ref{fig:dsrcphh0}, lens model. Compare Figure
  \ref{fig:dinvhsll0}.}

\plot{ddiffinvhsll0}{width=\textwidth}{Difference between image
  (Figure \ref{fig:predinvhslh0}) of $\Lambda^+_{z_s}$ applied to
  pressure gather (Figure \ref{fig:dsrcpph0}) and source gather
  (Figure \ref{fig:dinvhsll0}) inverted from receiver pressure gather
  (Figure \ref{fig:drecplh0}), lens model used in all propagations.
 Same color scale as in Figure \ref{fig:predinvhsll0}.}
